---
output: dfeR::official_statistic
params:
    title: "Blockbuster output"
    subtitle: "June 2018"
    path: "output2018-06-09.xlsx"
    forecast_horizon: 10
    block_rebuild_cost: 2000
    repair_order: "EDCB"
    det_rates: "default"
    repair_costs : "default"
    inflation: "yes"
    repair_money : "0, 0, 0, 0, 0, 0, 0, 0, 0, 0"
    rebuild_money: "0, 0, 0, 0, 0, 0, 0, 0, 0, 0"
always_allow_html: yes 
title: "`r params$title`"
subtitle: "`r params$subtitle`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# It is recommended that you load in any scripts your require in the markdown here. E.g:
library(ggplot2)
library(xlsx)
library(dplyr)
library(govstyle)
library(scales)

format_money <- function(m, digits = 2){
  if(length(digits) != 1 && length(m) != length(digits)) stop("The two arguments must be of the same length.")
  neg <- m < 0 & !is.na(m) # to identify negative values (NAs are false)
  m <- m * (-1)^ neg
  exponent <- floor(log10(m) / 3)
  exponent[exponent < 0] <- 0
  suffix <- c("", "K", "M", "B", "T")[exponent + 1]
  results <- paste0("£", signif(m, digits) / (10 ^ (exponent * 3)), suffix)
  results[is.na(m)] <- NA
  results[neg] <- paste0("-", results)
  return(results)
}
```

```{r data, echo = FALSE, warning = FALSE, message = FALSE}
backlog <- read.xlsx(file.path(params$path), 1) %>%
  mutate(`Year on year change` = backlog - c(NA, backlog[-length(backlog)]))
totals <- read.xlsx(file.path(params$path), 2)
```

# Model parameters

The following outputs are from a Blockbuster Deterioration model run over `r params$forecast_horizon` years. A summary of the model parameters in provided below, but full details can be found in the accompanying **Excel input.xlsm** file.

```{r params, results = "asis"}

knitr::kable(data.frame(Parameter = c(#"Deterioration rates", "Repair costs", 
                                      "Inflation on repair and rebuild costs",
                                      "Block rebuild unit cost", "Repair order"),
                        "Value" = c(#params$det_rates, params$repair_costs,
                               params$inflation, params$block_rebuild_cost,
                               paste(params$repair_order, collapse = ", "))))

# repair_money <- inputs$repair_budget
# rebuild_money <- inputs$rebuild_budget
repair_money <- strsplit(params$repair_money, ", ")[[1]] %>% as.numeric
rebuild_money <- strsplit(params$rebuild_money, ", ")[[1]] %>% as.numeric
```

##

The yearly budgets available for rebuilding and repairing are:

```{r budgets}

knitr::kable(data.frame(Year = seq_len(params$forecast_horizon),
                        `Repair budget` = format_money(repair_money),
                        `Rebuild budget` = format_money(rebuild_money)))

```

# Model results

## Expected backlog per year

```{r backlog_table, fig.align = "center", fig.cap = "Yearly backlog forecasts"}
knitr::kable(backlog %>%
               mutate(`Year on year change` = format_money(`Year on year change`),
                      backlog = format_money(backlog)),
             row.names = FALSE,
             col.names = c("Year", "Backlog", "Year on year change"))
```

## Expected backlog per year

```{r backlog, echo = FALSE, warning = FALSE}
backlog %>%
  ggplot(aes(x = factor(year), y = backlog, group = 1)) + 
  geom_line() + 
  theme_gov() +
  xlab("Year") +
  ylab("Backlog") +
  scale_y_continuous(label = format_money)
```

## Expected backlog by condition grades

```{r stacked_backlog, echo = FALSE, warning = FALSE}
axis_breaks <- seq(0, max(totals$year), max(totals$year) / 10)
totals %>%
  filter(grade != "A") %>%
  mutate(grade = case_when(       # combine grades D and E as E is model internal
    grade %in% c("D", "E") ~ "D",
    TRUE                   ~ as.character(grade))) %>%
  group_by(grade, year) %>%
  summarise(backlog = sum(backlog),
            area = sum(area)) %>%
  ggplot(aes(x = year, backlog, fill = grade)) + 
  geom_area() + 
  theme_gov() +
  xlab("Year") +
  ylab("Backlog") + 
  theme(legend.position = "top") +
  scale_y_continuous(labels=format_money) +
  scale_x_continuous(labels = as.integer, breaks = axis_breaks)
```



## Yearly change in expected backlog

```{r yearonyear, warning = FALSE}
backlog %>%
  ggplot(aes(x = factor(year), y = `Year on year change`, group = 1)) + 
  geom_line() + 
  theme_gov() +
  xlab("Year") +
  scale_y_continuous(label = format_money) +
  expand_limits(y = 0)
```

## Expected proportion by condition grades

```{r area, echo = FALSE, warning = FALSE}
total_area <- totals %>% filter(year == 0) %>% pull(area) %>% sum

totals %>%
  mutate(grade = case_when(       # combine grades D and E as E is model internal
    grade %in% c("D", "E") ~ "D",
    TRUE                   ~ as.character(grade))) %>%
  group_by(grade, year) %>%
  summarise(backlog = sum(backlog),
            area = sum(area)) %>%
  ggplot(aes(x = year, area/total_area, fill = grade)) + 
  geom_area() + 
  theme_gov() +
  xlab("Year") +
  ylab("Percentage of estate") + 
  theme(legend.position = "top") +
  scale_y_continuous(labels=scales::percent) +
  scale_x_continuous(breaks = axis_breaks)
```
