---
output: dfeR::official_statistic
params:
    title: "Blockbuster output"
    subtitle: "June 2018"
    path: "output2018-06-09.xlsx"
    forecast_horizon: 10
    block_rebuild_cost: 2000
    repair_order: "EDCB"
    det_rates: "default"
    repair_costs : "default"
    inflation: "yes"
    repair_money : "0, 0, 0, 0, 0, 0, 0, 0, 0, 0"
    rebuild_money: "0, 0, 0, 0, 0, 0, 0, 0, 0, 0"
always_allow_html: yes 
title: "`r params$title`"
subtitle: "`r params$subtitle`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# It is recommended that you load in any scripts your require in the markdown here. E.g:
library(ggplot2)
library(xlsx)
library(dplyr)
library(govstyle)
library(scales)

format_money <- function(m, digits = 2){
  if(length(digits) != 1 && length(m) != length(digits)) stop("The two arguments must be of the same length.")
  neg <- m < 0 & !is.na(m) # to identify negative values (NAs are false)
  m <- m * (-1)^ neg
  exponent <- floor(log10(m) / 3)
  exponent[exponent < 0] <- 0
  suffix <- c("", "K", "M", "B", "T")[exponent + 1]
  results <- paste0("£", signif(m, digits) / (10 ^ (exponent * 3)), suffix)
  results[is.na(m)] <- NA
  results[neg] <- paste0("-", results)
  return(results)
}
```

```{r data, echo = FALSE, warning = FALSE, message = FALSE}

start_year <- 2020

backlog <- read.xlsx(file.path(params$path), 1) %>%
  mutate(`Year on year change` = backlog - c(NA, backlog[-length(backlog)]),
         year = year + start_year)
totals <- read.xlsx(file.path(params$path), 2) %>%
  mutate(year = year + start_year)
```

# Output matrix

```{r}
repair_money <- strsplit(params$repair_money, ", ")[[1]] %>% as.numeric
rebuild_money <- strsplit(params$rebuild_money, ", ")[[1]] %>% as.numeric

last_year <- totals %>% filter(year == max(year)) %>%
   mutate(grade = case_when(       # combine grades D and E as E is model internal
    grade %in% c("D", "E") ~ "D",
    TRUE                   ~ as.character(grade)),
    grade = factor(grade, levels = c("D", "C", "B", "A"))) %>%
  group_by(grade) %>%
  summarise(area = sum(area), backlog = sum(backlog))
total_area <- sum(last_year$area)
last_year <- last_year %>% mutate(area = area / total_area)

`Level of investment` <- format_money(sum(rebuild_money + repair_money))
`Value of backlog` <- backlog %>% filter(year == max(year)) %>% pull(backlog) %>% format_money
`Number of schools rebuilt` <- NA
`Number of schools in need of rebuilding` <- NA
`Estate at A` <- last_year %>% filter(grade == "A") %>%
  pull(area)
`Estate at B` <- last_year %>% filter(grade == "B") %>%
  pull(area)
`Estate at C` <- last_year %>% filter(grade == "C") %>%
  pull(area)
`Estate at D` <- last_year %>% filter(grade == "D") %>%
  pull(area)

pupils <-  c(8406181, 8520189, 8605301, 8677929, 8738575, 8771860, 8789721,
             8784848, 8781830, 8766692, 8736775, 8689746, 8658017, 8635628,
             8613873, 8588705, 8578881, 8561901, 8546396, 8533374, 8524127,
             8519215, 8537072, 8556856, 8583943, 8617228, 8655872, 8698621,
             8744151, 8791072, 8837951, 8883329, 8930701, 8976555, 9019528,
             9058499, 9092502)
pupils <- data.frame(year = 2020:2056, pupils) %>%
  filter(year == max(backlog$year)) %>%
  pull(pupils)

`Pupils in C` <- pupils * `Estate at C`
`Pupils in D` <- pupils * `Estate at D`
`Pupils in A or B` <- pupils * (`Estate at A` + `Estate at B`)

`Estate at A` <- paste(round(`Estate at A` * 100, 1), "%", sep = "")
`Estate at B` <- paste(round(`Estate at B` * 100, 1), "%", sep = "")
`Estate at C` <- paste(round(`Estate at C` * 100, 1), "%", sep = "")
`Estate at D` <- paste(round(`Estate at D` * 100, 1), "%", sep = "")

output_matrix <- data.frame(`Level of investment`, `Value of backlog`,
                            `Number of schools rebuilt`, `Number of schools rebuilt`,
                            `Estate at A`,
                            `Estate at B`,
                            `Estate at C`,
                            `Estate at D`,
                            `Pupils in A or B`, `Pupils in C`, `Pupils in D`)

```


```{r results = "asis"}

knitr::kable(output_matrix, format = "markdown", format.args = list(big.mark = ","),
             col.names = c("Level of investment", "Value of backlog",
                            "Number of schools rebuilt", "Number of schools in need of rebuilding",
                            "Estate at A",
                            "Estate at B",
                            "Estate at C",
                            "Estate at D",
                            "Pupils in A or B", "Pupils in C", "Pupils in D"))

```

# Model parameters

The following outputs are from a Blockbuster Deterioration model run over `r params$forecast_horizon` years. A summary of the model parameters in provided below, but full details can be found in the accompanying **Excel input.xlsm** file.

```{r params, results = "asis"}

knitr::kable(data.frame(Parameter = c(#"Deterioration rates", "Repair costs", 
                                      "Inflation on repair and rebuild costs",
                                      "Block rebuild unit cost", "Repair order"),
                        "Value" = c(#params$det_rates, params$repair_costs,
                               params$inflation,
                               paste0("£", round(params$block_rebuild_cost,2), sep = ""),
                               paste(params$repair_order, collapse = ", "))),
             format = "markdown")

# repair_money <- inputs$repair_budget
# rebuild_money <- inputs$rebuild_budget

```

##

The yearly budgets available for rebuilding and repairing are:

```{r budgets}

knitr::kable(format = "markdown", data.frame(Year = seq_len(params$forecast_horizon) + start_year,
                        `Repair budget` = format_money(repair_money),
                        `Rebuild budget` = format_money(rebuild_money)))

```

# Model results

## Expected backlog per year

```{r backlog_table, fig.align = "center", fig.cap = "Yearly backlog forecasts"}
knitr::kable(format = "markdown", backlog %>%
               mutate(`Year on year change` = format_money(`Year on year change`),
                      backlog = format_money(backlog)),
             row.names = FALSE,
             col.names = c("Year", "Backlog", "Year on year change"))
```

## Expected backlog per year

```{r backlog, echo = FALSE, warning = FALSE, fig.height=4.5, fig.width = 6}
backlog %>%
  ggplot(aes(x = factor(year), y = backlog, group = 1)) + 
  geom_hline(yintercept = 1:7 * 1e10, linetype = "dotted", alpha = 0.3) +
  geom_bar(stat= "identity", fill = "#2B8CC4") + 
  theme_gov() +
  xlab("Year") +
  ylab("Backlog") +
  scale_y_continuous(label = format_money)
```

## Expected backlog by condition grades

```{r stacked_backlog, echo = FALSE, warning = FALSE, fig.height=4.5, fig.width = 6}
axis_breaks <- seq(0, max(totals$year), max(totals$year) / 10)
totals %>%
  filter(grade != "A") %>%
  mutate(grade = case_when(       # combine grades D and E as E is model internal
    grade %in% c("D", "E") ~ "D",
    TRUE                   ~ as.character(grade)),
    grade = factor(grade, c("D", "C", "B"))) %>%
  group_by(grade, year) %>%
  summarise(backlog = sum(backlog),
            area = sum(area)) %>%
  mutate(label = paste(grade, ": ", format_money(backlog), sep = "")) %>%
  ggplot(aes(x = factor(year), backlog, fill = grade)) + 
  geom_hline(yintercept = 1:7 * 1e10, linetype = "dotted", alpha = 0.3) +
  geom_bar(stat = "identity") + 
  theme_gov() +
  xlab("Year") +
  ylab("Backlog") + 
  scale_y_continuous(labels=format_money) +
  scale_fill_manual(values = gov_cols[c("red", "yellow", "grass_green")] %>% unname) +
  geom_text(aes(label = label), vjust = 1, position = "stack")
  
  
  # scale_x_discrete(labels = as.integer, breaks = axis_breaks)
```



## Yearly change in expected backlog

```{r yearonyear, warning = FALSE, fig.height=4.5, fig.width = 6}
backlog %>%
  ggplot(aes(x = factor(year), y = `Year on year change`, group = 1)) + 
  geom_hline(yintercept = -0:7 * 1e9, linetype = "dashed", alpha = 0.3) +
  geom_line() + 
  geom_point() +
  geom_text(aes(label = year), vjust = 1, nudge_y = -1e8) + 
  theme_gov() +
  xlab("Year") +
  scale_y_continuous(label = format_money) +
  expand_limits(y = 0)
```

## Expected proportion by condition grades

```{r area, echo = FALSE, warning = FALSE, fig.height=4.5, fig.width = 6}
total_area <- totals %>% group_by(year) %>% summarise(total_area = sum(area))

totals %>%
  mutate(grade = case_when(       # combine grades D and E as E is model internal
    grade %in% c("D", "E") ~ "D",
    TRUE                   ~ as.character(grade)),
    grade = factor(grade, levels = c("D", "C", "B", "A"))) %>%
  left_join(total_area) %>%
  group_by(grade, year) %>%
  summarise(area = sum(area / total_area)) %>%
  mutate(label = paste(grade, ": ", round(area * 100,1), "%", sep = "")) %>%
  ggplot(aes(x = factor(year), y = area, fill = grade)) + 
  geom_bar(stat="identity") + 
  geom_text(aes(label = label), vjust = 1, position = "stack") +
  theme_gov() +
  xlab("Year") +
  ylab("Percentage of estate") + 
  scale_y_continuous(labels=scales::percent) +
  scale_fill_manual(values = gov_cols[c("red", "yellow", "grass_green", "green")] %>% unname)
```
